\thetitle

### 2013-04-17 tfpie/tfpie.tex
No % convention.

To begin with, we introduce the notion of a \emph{class definition},
which can be thought of at first as a structure definition in that
it defines a new class of compound data.  A class is defined using the
{\tt define-class} form:
\begin{verbatim}
    (define-class posn% (fields x y))
\end{verbatim}
An \emph{object} is a value that is a member of this class of data,
which can be constructed with the {\tt new} keyword, a class name, and
the appropriate number of arguments for the fields of the object:
\begin{verbatim}
    (new posn% 3 4)
\end{verbatim}
An object understands some set of \emph{messages}.  Simple
structure-like objects understand messages for accessing their fields
and message are sent by using the {\tt send} keyword, followed by an object,
a message name, and some number of arguments:
\begin{verbatim}
    (send (new posn% 3 4) x) ;=> 3
    (send (new posn% 3 4) y) ;=> 4
\end{verbatim}
The {\tt send} notation is simple, but syntactically heavy.  Once
students are comfortable with the {\tt send} form, we introduce a
shorthand to make it more convenient by writing {\tt (x . m)} for {\tt
  (send x m)}.  The \emph{dot notation} can be nested, so {\tt (x . m
  . n)} is shorthand for {\tt (send (send x m) n)}.  (The approach of
introducing a simple, uniform syntax and later introducing a
convenient shorthand that would have been confusing to start with
follows the approach of introducing {\tt cons} and later {\tt list}
and {\tt quote} in the first semester.)

It is possible to endow objects with functionality by defining \emph{methods},
which extend the set of messages an object understands.  A method definition
follows the same syntax as a function definition, but is located inside of a
class definition.  Here is a more complete development of the {\tt posn\%} class
that includes a couple of methods:
\begin{verbatim}
    ;; A Posn is a (new posn% Number Number),
    ;; which represents a point on the Cartesian plane
    (define-class posn% (fields x y)

      ;; Posn -> Number
      ;; Distance between this posn and that posn
      (check-expect ((new posn% 0 0) . dist (new posn% 3 4)) 5)
      (define (dist that)
        (sqrt (+ (sqr (- (this . x) (that . x)))
                 (sqr (- (this . y) (that . y))))))

      ;; -> Number
      ;; Distance of this posn from the origin
      (check-expect ((new posn% 0 0) . dist-origin) 0)
      (check-expect ((new posn% 3 4) . dist-origin) 5)
      (define (dist-origin)
        (this . dist (new posn% 0 0))))
\end{verbatim}

This class definition defines a new class of values which are {\tt
  posn\%} objects.  Such objects are comprised of two numeric values
and understand the messages {\tt x}, {\tt y}, {\tt dist}, and {\tt
  dist-origin}.  Unit tests have been included with each method
definition, following the principles of the design recipe studied in
the first semester.  In fact, the {\tt check-expect} mechanism works
exactly as it did before.

@@

(check-expect (sum (make-leaf 7)) 7)
(check-expect 
  (sum (make-node
         (make-leaf 1)
         5 
         (make-node (make-leaf 0)
                    10
                    (make-leaf 0))))
  16)
\end{verbatim}
\end{minipage}
\begin{minipage}[t]{3in}
\begin{verbatim}
#lang class/1
;; A Tree is one of:
;; - (new leaf% Number)
;; - (new node% Tree Number Tree)
;; and implements
;; sum : -> Number
;; sums the elements of this tree

(define-class leaf%
  (fields v)
  (define (sum) (this . v)))

(define-class node%
  (fields left v right)
  (define (sum)
    (+ (this . left . sum)
       (this . v)
       (this . right .sum))))

(check-expect ((new leaf% 7) . sum) 7)
(check-expect 
  ((new node%
        (new leaf% 1)
        5 
        (new node% (new leaf% 0) 
                   10
                   (new leaf% 0))))
    . sum)
  16)
\end{verbatim}
\end{minipage}
\caption{Binary tree sum in Beginning Student and in the Class language}
\label{fig:tree}
\end{figure}

The structure of this data definition is analagous to the approach of
the previous semester but this example brings to light an important
difference with the functional approach.  The method for computing the
sum of the empty tree is defined in the {\tt leaf\%} class, while the
method for computing the sum of a node is in the {\tt node\%} class.
When a tree object is sent the {\tt sum} method, there is no function
with a conditional to determine whether the object is a
leaf---instead, the object itself takes care of computing the sum
based on the {\tt sum} method it contains.  This shift in perspective
is at the core of object-orientation: objects contain their own
behavior and the case analysis previously done in functions is
eliminated.

@@

;; A World is a (new world% Number)
(define-class world%
  (fields n)

  ;; on-tick : -> World
  (define (on-tick)
    (new world% (add1 (this . n))))

@@

  ;; on-key : KeyEvent -> World
  (define (on-key k) (new world% 10)))

(big-bang (new world% 10))
\end{verbatim}
\end{minipage}
\caption{World programs}
\label{fig:world}
\end{figure}

>>>>

To begin with, we introduce the notion of a \emph{class definition},
which can be thought of at first as a structure definition in that
it defines a new class of compound data.  A class is defined using the
{\tt define-class} form:
\begin{verbatim}
    (define-class posn (fields x y))
\end{verbatim}
An \emph{object} is a value that is a member of this class of data,
which can be constructed with the {\tt new} keyword, a class name, and
the appropriate number of arguments for the fields of the object:
\begin{verbatim}
    (new posn 3 4)
\end{verbatim}
An object understands some set of \emph{messages}.  Simple
structure-like objects understand messages for accessing their fields
and message are sent by using the {\tt send} keyword, followed by an object,
a message name, and some number of arguments:
\begin{verbatim}
    (send (new posn 3 4) x) ;=> 3
    (send (new posn 3 4) y) ;=> 4
\end{verbatim}
The {\tt send} notation is simple, but syntactically heavy.  Once
students are comfortable with the {\tt send} form, we introduce a
shorthand to make it more convenient by writing {\tt (x . m)} for {\tt
  (send x m)}.  The \emph{dot notation} can be nested, so {\tt (x . m
  . n)} is shorthand for {\tt (send (send x m) n)}.  (The approach of
introducing a simple, uniform syntax and later introducing a
convenient shorthand that would have been confusing to start with
follows the approach of introducing {\tt cons} and later {\tt list}
and {\tt quote} in the first semester.)

It is possible to endow objects with functionality by defining \emph{methods},
which extend the set of messages an object understands.  A method definition
follows the same syntax as a function definition, but is located inside of a
class definition.  Here is a more complete development of the {\tt posn} class
that includes a couple of methods:
\begin{verbatim}
    ;; A Posn is a (new posn Number Number),
    ;; which represents a point on the Cartesian plane
    (define-class posn (fields x y)

      ;; Posn -> Number
      ;; Distance between this posn and that posn
      (check-expect ((new posn 0 0) . dist (new posn 3 4)) 5)
      (define (dist that)
        (sqrt (+ (sqr (- (this . x) (that . x)))
                 (sqr (- (this . y) (that . y))))))

      ;; -> Number
      ;; Distance of this posn from the origin
      (check-expect ((new posn 0 0) . dist-origin) 0)
      (check-expect ((new posn 3 4) . dist-origin) 5)
      (define (dist-origin)
        (this . dist (new posn 0 0))))
\end{verbatim}

This class definition defines a new class of values which are {\tt
  posn} objects.  Such objects are comprised of two numeric values
and understand the messages {\tt x}, {\tt y}, {\tt dist}, and {\tt
  dist-origin}.  Unit tests have been included with each method
definition, following the principles of the design recipe studied in
the first semester.  In fact, the {\tt check-expect} mechanism works
exactly as it did before.

@@

(check-expect (sum (make-leaf 7)) 7)
(check-expect 
  (sum (make-node
         (make-leaf 1)
         5 
         (make-node (make-leaf 0)
                    10
                    (make-leaf 0))))
  16)
\end{verbatim}
\end{minipage}
\begin{minipage}[t]{3in}
\begin{verbatim}
#lang class/1
;; A Tree is one of:
;; - (new leaf Number)
;; - (new node Tree Number Tree)
;; and implements
;; sum : -> Number
;; sums the elements of this tree

(define-class leaf
  (fields v)
  (define (sum) (this . v)))

(define-class node
  (fields left v right)
  (define (sum)
    (+ (this . left . sum)
       (this . v)
       (this . right .sum))))

(check-expect ((new leaf 7) . sum) 7)
(check-expect 
  ((new node
        (new leaf 1)
        5 
        (new node (new leaf 0) 
                  10
                  (new leaf 0))))
    . sum)
  16)
\end{verbatim}
\end{minipage}
\caption{Binary tree sum in Beginning Student and in the Class language}
\label{fig:tree}
\end{figure}

The structure of this data definition is analagous to the approach of
the previous semester but this example brings to light an important
difference with the functional approach.  The method for computing the
sum of the empty tree is defined in the {\tt leaf} class, while the
method for computing the sum of a node is in the {\tt node} class.
When a tree object is sent the {\tt sum} method, there is no function
with a conditional to determine whether the object is a
leaf---instead, the object itself takes care of computing the sum
based on the {\tt sum} method it contains.  This shift in perspective
is at the core of object-orientation: objects contain their own
behavior and the case analysis previously done in functions is
eliminated.

@@

;; A World is a (new world Number)
(define-class world
  (fields n)

  ;; on-tick : -> World
  (define (on-tick)
    (new world (add1 (this . n))))

@@

  ;; on-key : KeyEvent -> World
  (define (on-key k) (new world 10)))

(big-bang (new world 10))
\end{verbatim}
\end{minipage}
\caption{World programs}
\label{fig:world}
\end{figure}
